---
title: "Lab 2"
author: "Mason Witcraft"
date: "January 27, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

#### Working Directory
```{r}
getwd()
```

# Task 2

#### Read Data
```{r}
epa = read.csv("EPAGAS.csv")
head(epa)
```

# Task 3

#### Create *mpg* Vector
```{r}
mpg = epa$MPG
```
#### Define *z = ($x_i$ - $\bar{x}$)/$s_x$*
```{r}
z = (mpg - mean(mpg))/sd(mpg)
```
#### Verify *$\bar{z}$ = 0*
```{r}
mean(z)
```
#### Verify *$s^2_z$ = 1*
```{r}
var(z)
```
#### Find values of *mpg* that are possible outliers
```{r}
mpg[abs(z) >= 2 & abs(z) <= 3]
```
#### Find values of *mpg* that are defined as outliers
```{r}
mpg[abs(z) > 3]
```
#### Using the lattice package, contruct a dotplot with colors *Red=outlier* and *Blue=possible outlier*
```{r}
library(lattice)
colorOutliers = ifelse(abs(z) > 3, "Red", ifelse(abs(z) >= 2 & abs(z) <= 3, "Blue", "Black"))  
dotplot(mpg, col=colorOutliers)
```

# Task 4

#### Make a boxplot of the mpg variable
```{r}
with(epa, boxplot(MPG, ylab = "MPG", col = "green3", horizontal = TRUE))
title("MPG Boxplot")
```

#### Chebyshev's Theorem:
* 75% of the data must lie within 2 standard deviations of the mean
* Using the *mpg* data:
```{r}
length(mpg[abs(z) < 2])/length(mpg)
```
* Chebyshev does agree with this data, as 96% > 75%

#### Empirical Rule:
* 95% of the data must lie within 2 standard deviations of the mean
* Using the *mpg* data:
```{r}
length(mpg[abs(z) < 2])/length(mpg)
```
* This data does correspond well with the Empirical Rule
* The Empirical Rule is valid in this case because 96% of the data is within 2 standard deviations
* Also, the data is a standard bell curve, so the Empirical Rule is still valid in this case:
```{r}
hist(mpg)
```

